---
title: "p8105_hw3_wm2460"
author: "Wenyan Ma"
date: "2022-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)
```

## Problem 1

## Problem 2

### Step 1: Tidy the dataset
```{r tidy}
accel_df = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_prefix = "activity_", 
    names_to = "minute", 
    values_to = "activity") %>% 
  mutate(minute = as.numeric(minute), 
         day_of_week = recode(day,"Monday" = "Weekday", 
                                   "Tuesday" = "Weekday", 
                                   "Wednesday" = "Weekday", 
                                   "Thursday" = "Weekday", 
                                   "Friday" = "Weekday", 
                                   "Saturday" = "Weekend", 
                                   "Sunday" = "Weekend"))
```

#### Describe the dataset

In the `accel_df` dataset, there are `r nrow(accel_df)` observations and `r ncol(accel_df)` variables. The variables in this dataset include `r colnames(accel_df)`. 

### Step 2: Traditional analyses of accelerometer data
```{r traditional_analyses}
traditional = accel_df %>%
  group_by(week, day) %>%
  summarize(activity_sum = sum(activity)) %>% 
  pivot_wider(names_from = day,
              values_from = activity_sum)
```

#### Trend

The trend is apparent. The total activity across Monday to Friday is more than the total activity done during the weeekend. 


### Step 3: Single-panel plot
```{r single_plot}
accel_df %>% 
  ggplot(aes(x = minute/60, y = activity, color = day)) + 
  geom_line(alpha = .5) +
  geom_smooth(aes(group = day), se = FALSE) +
  labs(title = "24-hour activity time courses for each day",
       x = "Hour",
       y = "Activity measured by accelerometer") +
  theme_bw()

```

#### Patterns 

From the plot above, we can see that from hour 0 to hour 5, there were usually not a lot of activity detected by the accelerometer. Between hour 6 and hour 19, there are some more activity than the earlier hours. However, most intense activity patterns usually happen between hour 19 and hour 22. This pattern is quite consistent across all 7 days of the week. 


## Problem 3

```{r exploration}
data("ny_noaa")
```

### Exploration of this dataset
In the original `ny_noaa` dataset, there are `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. The variables in this dataset include `r colnames(ny_noaa)`.

There are a lot of observations with missing values in this dataset. For the variable `prcp`, there are `r sum(is.na(ny_noaa$prcp))` missing values. For the variable `snow`, there are `r sum(is.na(ny_noaa$snow))` missing values. For the variable `snwd`, there are `r sum(is.na(ny_noaa$snwd))` missing values. For the variable `tmax`, there are `r sum(is.na(ny_noaa$tmax))` missing values. For the variable `tmin`, there are `r sum(is.na(ny_noaa$tmin))` missing values. 

### Step 1: Data cleaning
```{r cleaning}
noaa_new = ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(tmax = as.numeric(tmax), 
         tmin = as.numeric(tmin),
         tmin = tmin / 10,
         tmax = tmax / 10) %>% 
  filter(prcp >= "0", snow >= "0")

```

```{r snowfall_mode}
ny_noaa %>% 
  count(snow) %>% 
  arrange(desc(n))
```

#### For snowfall, what are the most commonly observed values? Why?

For snowfall, the most commonly observed values is 0. Because it did not snow in most days in New York state. The second most observed value is `NA`, which means the value is missing. 

As a note, I believe that mm is an appropriate unit for the `snow` variable, therefore I will use the original variable to proceed. 


### Step 2: Two-panel plot for average max temp
```{r plot}
avg_max = noaa_new %>% 
  group_by(id, year, month) %>% 
  filter(month %in% c(1, 7)) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE, color = id)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id)) + 
    geom_point(alpha = 0.25) + 
    geom_line(alpha = 0.25) + 
    facet_grid(. ~ month) +
    labs(title = 
           "Average max temperature in January and in July in each station across years",
       x = "Year",
       y = "Average max temperature") 

avg_max
```

#### Is there any observable / interpretable structure? Any outliers?
In my opinion, it is clear that the average max temperature in January is much lower than the average max temperature in July. There were some outliers for both months, especially for January. There is one outlier for July as well. 


### Step 3: Plots for tmin vs tmax and snowfall

#### Plot for tmin vs tmax
```{r tmin_tmax}
vs_p = 
  noaa_new %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() +
  labs(title = "Tmin vs Tmax") 

vs_p

```


#### Plot for snowfall
```{r snowfall}
snowfall_p = 
  noaa_new %>% 
  filter(snow %in% (0:100)) %>% 
  group_by(year) %>% 
  ggplot(aes(x = snow, y = as.factor(year))) +
  geom_density_ridges() +
  labs(title = "Distribution of snowfall values by year", 
       x = "Snowfall in mm",
       y = "Year") 
  
snowfall_p
```

#### Making the plots into a two-panel plot
```{r combine_plots, width = 10}
(vs_p + snowfall_p)
```






