---
title: "p8105_hw3_wm2460"
author: "Wenyan Ma"
date: "2022-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

### Step 1: Tidy the dataset
```{r tidy}
accel_df = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_prefix = "activity_", 
    names_to = "minute", 
    values_to = "activity") %>% 
  mutate(minute = as.numeric(minute), 
         day_of_week = recode(day,"Monday" = "Weekday", 
                                   "Tuesday" = "Weekday", 
                                   "Wednesday" = "Weekday", 
                                   "Thursday" = "Weekday", 
                                   "Friday" = "Weekday", 
                                   "Saturday" = "Weekend", 
                                   "Sunday" = "Weekend"))
```

#### Describe the dataset

In the `accel_df` dataset, there are `r nrow(accel_df)` observations and `r ncol(accel_df)` variables. The variables in this dataset include `r colnames(accel_df)`. 

### Step 2: Traditional analyses of accelerometer data
```{r traditional_analyses}
traditional = accel_df %>%
  group_by(week, day) %>%
  summarize(activity_sum = sum(activity)) %>% 
  arrange(week, desc(activity_sum)) %>%
  pivot_wider(names_from = day,
              values_from = activity_sum) %>% 
  knitr::kable()

traditional
```

#### Trend

The trend is apparent. The total activity across Monday to Friday is more than the total activity done during the weeekend. 


### Step 3: Single-panel plot
```{r single_plot}
accel_df %>% 
  ggplot(aes(x = minute/60, y = activity, color = day)) + 
  geom_line(alpha = .5) +
  labs(title = "24-hour activity time courses for each day",
       x = "Hour",
       y = "Activity measured by accelerometer") +
  theme_bw()


# or add  geom_smooth(aes(group = day), se = FALSE) + before labs #
```

#### Patterns 

From the plot above, we can see that from hour 0 to hour 5, there were usually not a lot of activity detected by the accelerometer. Between hour 6 and hour 19, there are some more activity than the earlier hours. However, most intense activity patterns usually happen between hour 19 and hour 22. This pattern is quite consistent across all 7 days of the week. 


## Problem 3

```{r exploration}
data("ny_noaa")
```

### Exploration of this dataset
In the original `ny_noaa` dataset, there are `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. The variables in this dataset include `r colnames(ny_noaa)`.

There are a lot of observations with missing values in this dataset. For the variable `prcp`, there are `r sum(is.na(ny_noaa$prcp))` missing values. For the variable `snow`, there are `r sum(is.na(ny_noaa$snow))` missing values. For the variable `snwd`, there are `r sum(is.na(ny_noaa$snwd))` missing values. For the variable `tmax`, there are `r sum(is.na(ny_noaa$tmax))` missing values. For the variable `tmin`, there are `r sum(is.na(ny_noaa$tmin))` missing values. 

### Step 1: Data cleaning
```{r cleaning}
noaa_new = ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(tmax = as.numeric(tmax), 
         tmin = as.numeric(tmin),
         tmin = tmin / 10,
         tmax = tmax / 10,
         prcp = prcp / 10) 

```

```{r snowfall_mode}
ny_noaa %>% 
  count(snow) %>% 
  arrange(desc(n))
```

#### For snowfall, what are the most commonly observed values? Why?

For snowfall, the most commonly observed values is 0. Because it did not snow most days in New York state during the given timeframe. The second most observed value is `NA`, which means the value is missing. 

As a note, I believe that mm is an appropriate unit for the `snow` variable, therefore I will use the original variable to proceed. 


### Step 2: Two-panel plot for average max temp
```{r plot}
avg_max = noaa_new %>% 
  group_by(id, year, month) %>% 
  filter(month %in% c("1", "7")) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax)) + 
    geom_point(alpha = 0.25) + 
    geom_line(alpha = 0.25) + 
    facet_grid(. ~ month) +
    labs(title = 
           "Average max temperature in January and in July in each station across years",
       x = "Year",
       y = "Average max temperature") 

avg_max
```

#### Is there any observable / interpretable structure? Any outliers?
In my opinion, it is clear that the average max temperature in January is much lower than the average max temperature in July. There were some outliers for both months, especially for January. Also, there is one outlier for July as well. 


### Step 3: Plots for tmin vs tmax and snowfall

#### Plot for tmin vs tmax
```{r tmin_tmax}
vs_p = 
  noaa_new %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() +
  labs(title = "Tmin vs Tmax") 

vs_p

```


#### Plot for snowfall
```{r snowfall}
snowfall_p = 
  noaa_new %>% 
  group_by(year) %>% 
  filter(snow %in% (0:100)) %>% 
  ggplot(aes(x = snow, y = as.factor(year))) +
  geom_density_ridges() +
  labs(title = "Distribution of snowfall values by year", 
       x = "Snowfall in mm",
       y = "Year") 
  
snowfall_p
```

#### Making the plots into a two-panel plot
```{r combine_plots}
(vs_p + snowfall_p)
```


### End of HW3


